Paper

### 2020-01-21 Paper.Rmd
replacing the original Rmd file to avoid (my) confusion

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
library(tidyverse)
library(BarrettsProgressionRisk)

library(ggrepel)
#library(GGally)
library(kableExtra)
library(reshape2)
library(gridExtra)
#library(CoxHD)
library(RColorBrewer)

@@

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info

patient.info = all.patient.info %>% dplyr::filter(Set == 'Training')

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}

chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

arm.loc = arrange(rbind(
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])

locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = cbind.data.frame(locs, globalCNdf)
globalCNdf = melt(globalCNdf, id.vars=colnames(locs))

globalCNdf$seg.type = '5MB'
globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')

globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn value
globalCNdf = globalCNdf %>% dplyr::group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = 
globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')# + theme(text=element_text(size=12))
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)

@@

```{r loaddata, message=F, warning=F, echo=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

cx = dysplasia.df[,'cx']

x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

@@

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}
perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

```{r, echo=F, fig.width=6, fig.height=6}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

cuts = seq(0,1,0.1) 
pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, pg.sampNOHGD %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% dplyr::group_by(quant) %>% dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

noHGD.pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

noHGD.pred.confidence$Risk = 'Moderate'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))

noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

BarrettsProgressionRisk::showPredictionCalibration(noHGD.pred.confidence) + labs(subtitle='HGD excluded model')

# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

  fitNOHGD = models[[select.alpha]]

  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[intersect(hgd.samples$Samplename, rownames(dysplasia.df)),]

  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
  colnames(pred) = 'Prediction'
  hgd.samples = base::merge(hgd.samples, pred, by.x='Samplename', by.y='row.names')

  hgd.pred = hgd.samples %>% dplyr::group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

  rm(no.hgd.plots, coefs, performance.at.1se, models)
}
#hist(hgd.samples$Prediction)

#rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p

ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  mutate_at(vars(start, end), list(as.numeric)) %>% mutate(chr = factor(chr,levels=c(1:22)))

features = features %>% mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), 
                    gl = factor(ifelse(coef > 0, 'gain','loss')))

features = left_join(features, chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}

file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

@@

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = pg.samp %>% dplyr::select(Status, Prediction)
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
roc$model = 'CN Model'
roc$is = 'discovery'

@@

insilico.rocs[['discovery']] = roc

df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
)))

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1, color='lightblue') + 
    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

# Not sure why I did this
df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
    est=t(round(pROC::coords(roc, "best", transpose=T),2)),
    max=t(round(pROC::coords(roc, 0.8, transpose=T),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose=T),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6, transpose=T),2))
)
thresholds$label=rownames(thresholds)

@@

# roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
# pROC::coords(rocNOHGD, "best")
# pROC::coords(rocNOHGD, 0.6)
# pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)

cutoff = round(pROC::coords(roc, 'best', ret = 'threshold', transpose=T), 2)
```

@@

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
```

@@

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')

rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First'),
  bind_rows(
    pROC::coords(rocEndoM, 'best', transpose=T),
    pROC::coords(rocEndoA, 'best', transpose=T),
    pROC::coords(rocLastEndo, 'best', transpose=T),
    pROC::coords(rocFirstEndo, 'best', transpose=T)
  ))

@@

roc$model = "CN Model"

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)

@@

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

```{r echo=F, fig.height=8, fig.width=6}
roc$model = "Per Sample"

>>>>

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2019-11-13"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(BarrettsProgressionRisk)
library(ggrepel)
library(kableExtra)
library(reshape2)
library(gridExtra)
library(RColorBrewer)

@@

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
patient.info = all.patient.info

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=6, fig.width=10}
chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

arm.loc = arrange(bind_rows(
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

#ddf = rbind(dysplasia.df, 'mean.value'=apply(dysplasia.df, 2, mean))
globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(patient.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))

globalCN.plot = cn.mtn.plot(globalCNdf %>% filter(seg.type != 'arms'), labeller(Status=n.status, chr=label_value))

cnLegend = get.legend(globalCN.plot)
globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'NDBE')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'NDBE')$variable)),')'))

globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot +  labs(y='mean adj. seg.', title='All samples') + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)

gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)

@@

## *Complexity Scores*

```{r loaddata, message=F, warning=F, echo=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

cx = dysplasia.df[,'cx']
complexity.score = tibble('Status'=labels, 'cx'=cx[names(labels)], 'Samplename' = names(labels)) %>%
  mutate(Status = factor(Status, labels = c('Non-Progressor','Progressor'))) %>%
  left_join(patient.info[,c('Samplename','Pathology','Patient')],by=c('Samplename'))

tt = wilcox.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

@@

rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

@@

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
hazards = apply( exp( t(dysplasia.df[,rownames(coefs[[select.alpha]])]) * coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)

cvRR = features

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

cvRR %>% mutate(`Gain/Loss` = ifelse(coef > 0 ,'Gain', 'Loss')) %>% 
  dplyr::select(label, hazards, `Gain/Loss`) %>% arrange(desc(hazards)) %>%
  write_tsv('plots/coef_effect.tsv')

@@

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr,levels=c(1:22))) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

@@

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
# loo_residuals = do.call(bind_rows, purrr::map(list.files(paste0(dir, '/Analysis/pcf_perPatient/50kb'), 'residuals.tsv', recursive = T, full.names = T), function(f) {
#   read_tsv(f, col_types = 'ccddddl')
# }))
# 
# failedQC = loo_residuals %>% filter(varMAD_median > 0.008)

error = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/sample_error.tsv')

file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

error = error %>% left_join(pg.samp %>% dplyr::select(Status, Samplename, RR), by=c('Sample'='Samplename')) %>% 
  dplyr::mutate(P.High = BarrettsProgressionRisk:::pi.hat(RR+Error), P.Low = BarrettsProgressionRisk:::pi.hat(RR-Error), Abs.Error = P.High-P.Low )

error %>% group_by(Status) %>% dplyr::summarise_at(vars(Abs.Error), list(~mean, ~sd))
summary(error$Abs.Error)
sd(error$Abs.Error)

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))

pg.samp = pg.samp %>% left_join(error %>% dplyr::select(-RR,-Status), by=c('Samplename'='Sample'))

@@

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=F)
```

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = pg.samp %>% dplyr::select(Status, Prediction)
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')

pROC::ci.auc(roc$auc)

ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)
ggsave(filename='plots/auc/training_auc_supp.png', plot=roc.plot(roc) + ggtitle('Per sample'), width=6, height=6, units='in', dpi=300)

roc$model = 'CN Model'
roc$is = 'discovery'

@@

insilico.rocs[['discovery']] = roc
df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('AUC' = as.numeric(r$auc), 'Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
)))

i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best') %>% as_tibble %>% mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))

#summary(filter(df, model == 'in-silico')$AUC)

getPalette = colorRampPalette(brewer.pal(5, "Blues"))
p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
  geom_line(linetype='solid', size=0.5) + 
  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
  scale_x_reverse() +
  scale_color_manual(values=getPalette(length(unique(df$is)))) +
  labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
p

blues = RColorBrewer::brewer.pal(5,'Blues')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

# Not sure why I did this
df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

@@

cutoff = round(pROC::coords(roc, 'best', ret = 'threshold'), 2)
```

@@

pROC::ci.auc(p53roc$auc)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
```

@@

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

pendo.rand = pg.samp %>% dplyr::group_by(Patient, Status, PID) %>%
  dplyr::summarise( r.Prob = sample(Prediction, 1), n = length(PID) ) %>% filter(n > 1)
rocRandEndo = pROC::roc(Status ~ r.Prob, data=pendo.rand, ci=T, of='thresholds')
rocRandEndo$model = 'Per Endoscopy Random'
ggsave('plots/auc/randomendo_auc.png', plot=roc.plot(rocRandEndo, title='Random per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First','Random'),
  bind_rows(
    pROC::coords(rocEndoM, 'best'),
    pROC::coords(rocEndoA, 'best'),
    pROC::coords(rocLastEndo, 'best'),
    pROC::coords(rocFirstEndo, 'best'),
    pROC::coords(rocRandEndo, 'best')
  ))
```

@@

roc$model = "CN Model"
m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)

@@

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_y_continuous(limits=c(0,100)) +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3, height=4, units="in", limitsize=F, dpi = 300)

@@

```{r}
roc$model = "Per Sample"
